```{r load_requirements, eval = T, include = F}  
# this loads the befdata package
require("rbefdata")  
require("ggplot2") 
```

## Abstract 

* Lots of data available
* Many accessible (with some hindrances) 
    - metadata missing (renders data useless)
* Many get still lost (needs help e.g rebind)
* Data often hard to find/use 

## Introduction 

Over the last decade ecology has grown into a more collaborative and
interdisciplinary science, to address questions on a greater temporal and
spatial scale (e.g michener et al 2012). With a growing awareness about the
value of data in the long run many efforts focused on building data platforms
ranging from small scales, covering a single project or collaboration, up to
global data bases and data warehouses (citation xx).  As it is harder to reuse
data without contextual information, many effort was also put into building up
standard frameworks that help to describe data like the Ecological Metadata
Language (cite xxx, other standards?). 

On characteristic of the data available in Biodiversity Ecosystem Functioning
is high homogeneity and large dataset size. It is generated partly automated by
remote sensing projects (e.g. Collins2006) using sensor or satellite networks,
enabling researchers to gain almost real-time insight into measured processes
(e.g michener ecoinformatics 2012}. And on the other hand there is data
characterized by a relative small dataset size and a high amount of
heterogeneity which gets generated by small studies spread all over the world
(e.g heidorn2009shedding).

## Material and Methods 

### BEFdata portal

The BEFdata portal is a data management platform developed within the 
BEF-China project (FOR xxx) of the german science foundation. It offers 
tools for data harmonization...

### Data used

The data used for the presentation of this package stems from ...

### Rbefdata

The `rbefdata` package offers an option list that is used to determine 
the servers URLs the package contacts to to retrieve data, the download 
forlder name for attached freeformat content of a dataset and 

* load the package

```{r require_rbefdata}
require(rbefdata)
```

* list options

```{r list_rbefdata_options}
bef.options()
```

* query options

```{r query_rbefdata_options}
bef.options("url")
```

* set options 

```{r set_rbefdata_options, eval = F}
bef.options("user_credentials" = "aölkjspoiul12")
bef.options("url" = "http://my.own.befdat.instance.com")
```

```{r load_secret_credentials, eval = T, include = F}  
# load your credentials
source("./secrets.R") 
```

Introduction to the rbefdata R package

* get datasets 

In the very heart of the BEFdata portal there is a paper proposal process integrated. You shop together 
datasets and afterwards create a paper proposal based on the shopped dataset. In the proposal
you have to give information like a title for the proposal and a rationale describing how you 
intend to use the data and where and when to publish the results. If the proposal is handed in 
the authors will be informed that somebody likes to access their datasets and they can decide 
if they like to participate and how. After all authors have granted access on is good to go 
with the `rbefdata` package. One can draw all datasets associated to a proposal in one turn 
with the package.

```{r rbefdata_get_datasets_from_proposal}
dataset_list = bef.portal.get.datasets_for_proposal(id = 1)
extract_one_dataset = dataset_list[[1]]
```

* Inspect datasets

The BEFdata portal offers metadat in Ecological Metadata Language format standard for download. 
This is used in the `rbefdata` package to associate a downloaded dataset with informations that
are required to understand a dataset. This information can be extracted from a dataset with 
the R command `attributes()`

```{r rbefdata_metadata}
metadata = bef.portal.get.dataset_for_proposal(id = 1)
```

* vizualize the portal (keywords) 

see figure ... in appendix

## Results 

## Discussion

## Appendix

### Figures 

```{r vizalize_keywords, cache = T} 
bef.portal.vizualize.keywords()
```
### Tables
